---
title: "Exploration"
format: html
editor: visual
---

```{r, include = FALSE}
library(tidyverse)
library(ggplot2)
```

## Load in the Data

```{r}
load("../../fluanalysis/data/clean_symptoms.rds")
```

## Dora the Explorer!

**Goals:** - For each (important) variable, produce and print some numerical output (e.g. a table or some summary statistics numbers).

-   For each (important) continuous variable, create a histogram or density plot.

-   Create scatterplots or boxplots or similar plots for the variable you decided is your main outcome of interest and the most important (or all depending on number of variables) independent variables/predictors. For this dataset, you can pick and choose a few predictor variables.

-   Any other exploration steps that might be useful.

### Continuous Variable: Body Temperature

Let's jump in with our continuous variable first.

**Summary Table**

```{r}
select_sympact%>%   #need to include na.rm exception since we have those 5 pesky NAs
  summarize(Avg = mean(BodyTemp, na.rm= T),
            Min = min(BodyTemp, na.rm= T),
            Max = max(BodyTemp, na.rm= T),
            Range = Max-Min,
            Q25 = quantile(BodyTemp, probs = .25, na.rm= T), 
            Median = median(BodyTemp, na.rm= T),
            Q75 = quantile(BodyTemp, probs = .75, na.rm= T))

```

Off the bat, the good news is it doesn't seem like we have any crazy outliers, as evidenced by the similarities between median and mean. Let's make a graph to see if the 103 is an "outlier" (not large enough to be a true outlier, but maybe a lone wolf) or if there's several participants with a fever \> 100.

Since we removed the unique visit number, we only have the one continuous variable, so this limits our possible EDA graph options.

**EDA Graphs: Body Temperature**

```{r, , warning=FALSE}
ggplot()+
  geom_bar(aes(x=BodyTemp), data=select_sympact, fill ="#0072B2")+    #I made it a pretty color for fun :)
  theme_bw()

```

So we can see that having a temperature over 100 isn't unusual. This is good to know moving forward. Are any of them considered true outliers?

```{r, warning=FALSE}
ggplot()+
  geom_boxplot(aes(BodyTemp), data=select_sympact, fill ="#56B4E9")+    #I made it a pretty color for fun :)
  theme_bw()

```

Alright this is good to know, even though a good number of people have temperatures greater than 101, by the spread available they are outliers in the data. Since there are so many and on a continuous scale, we won't remove them, but these bad boys may be the reason we see the (slight) difference between median and mean.

Off the top of my head when looking at possible predictors of body temperature, I would assume things like swollen lymph nodes, chills, subjective fever, myalgia, and weakness (Y/N) would be predictors. Let's do some quick explorations with these variables.

```{r}
select_sympact$Temp_group<-ifelse(between(select_sympact$BodyTemp , 97, 97.9), 97, 
                                  ifelse(between(select_sympact$BodyTemp , 98, 98.9), 98,
                                         ifelse(between(select_sympact$BodyTemp , 99, 99.9), 99,
                                                ifelse(between(select_sympact$BodyTemp , 100, 100.9), 100,
                                                       ifelse(between(select_sympact$BodyTemp , 101, 101.9), 101,
                                                              ifelse(between(select_sympact$BodyTemp , 102, 102.9), 102, 103))))))

table(select_sympact$Temp_group, select_sympact$SwollenLymphNodes)
table(select_sympact$Temp_group, select_sympact$ChillsSweats)
table(select_sympact$Temp_group, select_sympact$SubjectiveFever)
table(select_sympact$Temp_group, select_sympact$Weakness)
table(select_sympact$Temp_group, select_sympact$Myalgia)

```

Based on the numbers alone, it seems like all of these variables are strong indicators of body temperature. The most questionable would be swollen lymph nodes. Checking these out in the analysis will definitely be helpful and should provide some solid models.

Moving on...

Let's see what our categorical variable has in store for us.

### Categorical Variable: Nausea

First, let's see what our options for nausea entail:

```{r}
unique(select_sympact$Nausea)

```

Just, yes and no, no maybe. This is good to know.

Historically, for "extreme" categories with no midline, I've used JMP to help with analysis, especially with factorial models. So this will be a learning experience for me. Off the bat, let's compare general counts between the two options.

**Summary Table**

```{r}
select_sympact%>%   
  count(Nausea)
```

**EDA Graphs: Nausea**

```{r}
ggplot()+
  geom_bar(aes(x=Nausea), data=select_sympact, fill ="#0072B2")+    
  theme_bw()

```

For the benefit of the subjects, it's nice to see most people aren't experiencing nausea, though 258 cases isn't something to ignore.

Based on the symptoms available, I would assume that ab(domen) pain, chest pain, insomnia, vision, and vomit would be associated with nausea.

```{r}
table(select_sympact$Nausea, select_sympact$AbPain)
table(select_sympact$Nausea, select_sympact$ChestPain)
table(select_sympact$Nausea, select_sympact$Insomnia)
table(select_sympact$Nausea, select_sympact$Vision)
table(select_sympact$Nausea, select_sympact$Vomit)
```

Based on these contingency tables, the strongest evidence of a symptom of nausea is insomnia, but these will be further evaluated in the model as well.

### Both?

Great, now we have a basic understanding of these symptoms independently, but can we look at them together or if they relate to one another?

Let's make a basic contingency table with temp grouped by each degree

```{r}
table(select_sympact$Temp_group, select_sympact$Nausea)
```

I couldn't get the contingency table to give me conditional rates, so below are the calculated rates (%) of nausea among each temperature group:

**97**: 27.2

**98**: 36.4

**99**: 33.8

**100**: 38.9

**101**: 40.7

**102**: 18.2

**103**: 66.7

```{r, warning=FALSE}
ggplot()+
  geom_bar(aes(x=BodyTemp, fill = Nausea), position = "fill", data=select_sympact)+    
  theme_bw()

```

I chose a "fill" position here because since there was such a dense amount of body temperatures around 98 degrees it was hard to see if there was a trend between nausea and body temperature. Once we standardize the chart to represent the ratio at each degree (group) we can see there may be a general increase of rates of nausea with higher body temperatures, but we should measure this further (quantitatively).

When looking at the contingency tables, this is still a bit varied. The relationship may be something we look at more in the analysis, but for now, we have inconclusive results about the interrelation of nausea and body temperature.
